---
title: "F1_Data_Analysis"
author: "Miguel Ángel Ponce Martínez"
date: "18/2/2022"
output:
  html_document: 
    theme: cerulean
    highlight: tango
    number_sections: no
    toc: no
    toc_depth: 1
  pdf_document:
    theme: cerulean
    highlight: tango
    number_sections: yes
    toc: yes
    toc_depth: 1
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
```

#Libraries
```{r}
#These ones for Data Visualization
library(gridExtra)
library(ggplot2)
library(leaflet)
library(rgdal)
library(stringr)
library(reshape2)
library(forcats)
#These for PCA
library(tidyverse)
library(GGally) # ggplot2-based visualization of correlations
library(factoextra) # ggplot2-based visualization of pca
#Finally, those for Clustering
library(cluster)
library(mclust)

library(skimr)
library(mice)
library(VIM)
library(MASS)
library(glmnet)
library(e1071) 
library(rpart)
library(pROC)
library(class)
library(randomForest)
library(caret)
library(Rcpp)
library(naivebayes)
library(rpart.plot)
```

#Import Data
```{r}
setwd("C:/Users/migue/Documents/F1 Data Center/Datos")
data = read.csv("data.csv", sep= ';')
```

#Deal with NULL VALUES('NO TIME')
Assign NA's properly
```{r}
index = which(data$fp1_quali_Pace == 'NO TIME ') 
data$fp1_quali_Pace[index] = NA
index = which(data$fp1_race_Pace == 'NO TIME ') 
data$fp1_race_Pace[index] = NA
index = which(data$fp1_fastest_lap == 0) 
data$fp1_quali_Pace[index] = NA

index = which(data$fp2_quali_Pace == 'NO TIME ') 
data$fp2_quali_Pace[index] = NA
index = which(data$fp2_race_Pace == 'NO TIME ') 
data$fp2_race_Pace[index] = NA
index = which(data$fp2_fastest_lap == 0) 
data$fp2_quali_Pace[index] = NA

index = which(data$fp3_quali_Pace == 'NO TIME ') 
data$fp3_quali_Pace[index] = NA
index = which(data$fp3_race_Pace == 'NO TIME ') 
data$fp3_race_Pace[index] = NA
index = which(data$fp3_fastest_lap == 0) 
data$fp3_quali_Pace[index] = NA

index = which(data$quali_Pace == 'NO TIME ') 
data$quali_Pace[index] = NA
index = which(data$quali_fastest_lap == 0) 
data$quali_fastest_lap[index] = NA

index = which(data$race_Pace == 'NO TIME ') 
data$race_Pace[index] = NA
index = which(data$race_fastest_lap == 0) 
data$race_fastest_lap[index] = NA
```

#Data Preprocessing
```{r}
data = data[-which(data$quali_Laps == 0),]


aggr(data, numbers = TRUE, sortVars = TRUE, labels = names(data), cex.axis = 0.7, gap = 1.5, 
     ylab= c('Missing data','Pattern'))

#Remove those with over 0.6 of null values
data = data[,-c(24,25,34,36)]
```

```{r}
#Try to use mice library to make data imputation
#If it does not work, for example True or False about if the session started or not could be useful.
```

```{r}
#From now, I will keep 0 in 'NO TIME' values, if it does not perform, I could change it.
for (i in 1:ncol(data)){
  nul = which(is.na(data[,i]))
  data[c(nul),i] = 0
}
```

Factorize and numeric values
```{r}
for (i in 1:ncol(data)){
  if(i %in% c(8,10,15,17,22,27)) {
    data[,i] = as.integer(data[,i])
  } else {
    if(!(is.integer(data[,i]))) {
    data[,i] = as.factor(data[,i])}
  }
}
```

Correlations among data
```{r}
ggcorr(data, label = T,label_alpha=T,nbreaks = 4)
```

Teams Colors
```{r}
fer = '#9f121d'
rb = '#340db0'
mer = '#25b8a1'
am = '#185d5b'
at = '#2d3c4f'
ar = '#6b101b'
alp = '#2180c3'
wil = '#1e52b8'
mcl = '#ca923e'
haa = '#f0eff2'


colour_list = c('#ca923e', '#ca923e', '#185d5b', '#1e52b8', '#6b101b', '#f0eff2', '#2d3c4f', '#340db0', '#2180c3', '#9f121d', 
'#185d5b', '#2d3c4f', '#2180c3', '#340db0', '#25b8a1', '#f0eff2', '#9f121d', '#1e52b8', '#25b8a1', '#6b101b', '#6b101b')
```

Young vs Old drivers (old over 30 years old)
```{r}
#Incident rate
y = which(data$Age < 30)
young = data[y,]
old = data[-y,]

fp1_crash_old = length(which(old$fp1_Incident == 'Yes'))
fp1_crash_young = length(which(young$fp1_Incident == 'Yes'))
fp2_crash_old = length(which(old$fp2_Incident == 'Yes'))
fp2_crash_young = length(which(young$fp2_Incident == 'Yes'))
fp3_crash_old = length(which(old$fp3_Incident == 'Yes'))
fp3_crash_young = length(which(young$fp3_Incident == 'Yes'))
quali_crash_old = length(which(old$quali_Incident == 'Yes'))
quali_crash_young = length(which(young$quali_Incident == 'Yes'))
race_crash_old = length(which(old$race_Incident == 'Yes'))
race_crash_young = length(which(young$race_Incident == 'Yes'))

#Each session
y_o = c('Old_Drivers', 'Young_Drivers')
crash = c(fp1_crash_old,fp1_crash_young)
fp1 = data.frame(crash,y_o)
fp1 = ggplot(data=fp1, aes(x=reorder(y_o,crash), y=crash)) +
  geom_bar(stat="identity", fill="red")+
  geom_text(aes(label=crash), vjust=1.6, color="white", size=3.5)+
  theme_minimal()+labs(x = 'FP1')

crash = c(fp2_crash_old,fp2_crash_young)
fp2 = data.frame(crash,y_o)
fp2 = ggplot(data=fp2, aes(x=reorder(y_o,crash), y=crash)) +
  geom_bar(stat="identity", fill="red")+
  geom_text(aes(label=crash), vjust=1.6, color="white", size=3.5)+
  theme_minimal()+labs(x = 'FP2')

crash = c(fp3_crash_old,fp3_crash_young)
fp3 = data.frame(crash,y_o)
fp3 = ggplot(data=fp3, aes(x=reorder(y_o,crash), y=crash)) +
  geom_bar(stat="identity", fill="red")+
  geom_text(aes(label=crash), vjust=1.6, color="white", size=3.5)+
  theme_minimal()+labs(x = 'FP3')

crash = c(quali_crash_old,quali_crash_young)
quali = data.frame(crash,y_o)
quali = ggplot(data=quali, aes(x=reorder(y_o,crash), y=crash)) +
  geom_bar(stat="identity", fill="red")+
  geom_text(aes(label=crash), vjust=1.6, color="white", size=3.5)+
  theme_minimal()+labs(x = 'Qualyfing')

crash = c(race_crash_old,race_crash_young)
race = data.frame(crash,y_o)
race = ggplot(data=race, aes(x=reorder(y_o,crash), y=crash)) +
  geom_bar(stat="identity", fill="red")+
  geom_text(aes(label=crash), vjust=1.6, color="white", size=3.5)+
  theme_minimal()+labs(x = 'Race')

plot <- grid.arrange(fp1, fp2, fp3, quali, race, ncol=5)
```

As expected, young drivers crash more often than old drivers do.

Incident rate by driver
```{r}
driver_list = c('Daniel RICCIARDO', 'Lando NORRIS', 'Sebastian VETTEL', 'Nicholas LATIFI', 'Kimi RAIKKONEN', 'Nikita MAZEPIN', 'Pierre GASLY', 'Sergio PEREZ', 'Fernando ALONSO', 'Charles LECLERC', 'Lance STROLL', 'Yuki TSUNODA', 'Esteban OCON', 'Max VERSTAPPEN', 'Lewis HAMILTON', 'Mick SCHUMACHER', 'Carlos SAINZ', 'George RUSSELL', 'Valtteri BOTTAS', 'Antonio GIOVINAZZI','Robert KUBICA')

incidents <- vector()
for (i in 1:length(driver_list)){
fp1_crash = length(which(data$fp1_Incident == 'Yes' & data$Driver == driver_list[i]))
fp2_crash = length(which(data$fp2_Incident == 'Yes'& data$Driver == driver_list[i]))
fp3_crash = length(which(data$fp3_Incident == 'Yes'& data$Driver == driver_list[i]))
quali_crash = length(which(data$quali_Incident == 'Yes'& data$Driver == driver_list[i]))
race_crash = length(which(data$race_Incident == 'Yes'& data$Driver == driver_list[i]))
if (driver_list[i] == 'Max VERSTAPPEN' | driver_list[i] == 'Lewis HAMILTON'){
  if (driver_list[i] == 'Max VERSTAPPEN'){
    fp1_crash = fp1_crash + 1
  }else{
    fp1_crash = fp1_crash + 2
  }
}
total_crash = race_crash + quali_crash + fp1_crash +fp2_crash + fp3_crash
incidents[i] <- c(total_crash)
}

driv_short = c('RIC', 'NOR', 'VET', 'LAT', 'RAI', 'MAZ', 'GAS', 'PER', 'ALO', 'LEC', 'STR', 'TSU', 'OCO', 'VER', 'HAM', 'SCH', 'SAI', 'RUS', 'BOT', 'GIO', 'KUB')
inc_rate = data.frame(incidents,driv_short,colour_list)

ggplot(data=inc_rate, aes(x=reorder(driv_short,incidents), y=incidents)) +
  geom_bar(stat="identity", fill=colour_list)+
  geom_text(aes(label=incidents), vjust=1.6, color="grey", size=3.5)+
  theme_minimal()+labs(x = '', y = 'Incidents')
```

#Not official championships
#Data Creation
```{r}
circuits = as.character(unique(data$GP))

fp1_matrix <- seq(1:506)
fp2_matrix <- seq(1:506)
fp3_matrix <- seq(1:506)
quali_matrix <- seq(1:506)

dim(fp1_matrix) <- c(22,23)
dim(fp2_matrix) <- c(22,23)
dim(fp3_matrix) <- c(22,23)
dim(quali_matrix) <- c(22,23)

fp1_matrix[1,1] <- 'Driver/Circuit'
fp2_matrix[1,1] <- 'Driver/Circuit'
fp3_matrix[1,1] <- 'Driver/Circuit'
quali_matrix[1,1] <- 'Driver/Circuit'

#Add Driver and Circuit names
for (i in 1:length(driver_list)){
  fp1_matrix[i+1,1] <- driver_list[i]
  fp2_matrix[i+1,1] <- driver_list[i]
  fp3_matrix[i+1,1] <- driver_list[i]
  quali_matrix[i+1,1] <- driver_list[i]
}

for (j in 1:length(circuits)){
  fp1_matrix[1,j+1] <- circuits[j]
  fp2_matrix[1,j+1] <- circuits[j]
  fp3_matrix[1,j+1] <- circuits[j]
  quali_matrix[1,j+1] <- circuits[j]
}

for (i in 1:length(driver_list)){
  for (j in 1:length(circuits)){
    if (length(which(data$Driver == driver_list[i] & data$GP == circuits[j]))!= 0){
      m <- which(data$Driver == driver_list[i] & data$GP == circuits[j])
    fp1_matrix[i+1,j+1] <- data$fp1_fastest_lap[m]
    fp2_matrix[i+1,j+1] <- data$fp2_fastest_lap[m]
    fp3_matrix[i+1,j+1] <- data$fp3_fastest_lap[m]
    quali_matrix[i+1,j+1] <- data$quali_fastest_lap[m]
    }else{
    fp1_matrix[i+1,j+1] <- c(0) 
    fp2_matrix[i+1,j+1] <- c(0) 
    fp3_matrix[i+1,j+1] <- c(0) 
    quali_matrix[i+1,j+1] <- c(0)
    }
  }
}
#FP1/FP2/FP3/QUALY MATRIX
for (i in 1:ncol(quali_matrix)-1){
  print(paste(quali_matrix[,i+1],sep = "", collapse = ","))
}
```

#FP1 Representation
```{r}
#2021
points = c(47, 98, 41, 2, 9, 0, 175, 150, 60, 188, 21, 43, 42, 431, 346, 0, 170, 6, 376, 17, 0)
point_frame = data.frame(points,driv_short,colour_list)

fp1 <- ggplot(data=point_frame, aes(x=reorder(driv_short,points), y=points)) +
  geom_bar(stat="identity", fill=colour_list)+
  geom_text(aes(label=points), vjust=1.6, color="grey", size=3.5)+
  theme_minimal()+labs(x = 'FP1', y = 'Points')
```

#FP2 Representation
```{r}
#2021
points = c(40, 99, 34, 0, 5, 0, 132, 117, 138, 156, 49, 42, 129, 285, 349, 0, 123, 0, 305, 17, 0)
point_frame = data.frame(points,driv_short,colour_list)

fp2 <-ggplot(data=point_frame, aes(x=reorder(driv_short,points), y=points)) +
  geom_bar(stat="identity", fill=colour_list)+
  geom_text(aes(label=points), vjust=1.6, color="grey", size=3.5)+
  theme_minimal()+labs(x = 'FP2', y = 'Points')
```

#FP3 Representation
```{r}
#2021
points = c(23, 110, 10, 0, 18, 0, 143, 219, 62, 103, 18, 47, 37, 380, 292, 0, 143, 2, 201, 10, 0)
point_frame = data.frame(points,driv_short,colour_list)

fp3 <-ggplot(data=point_frame, aes(x=reorder(driv_short,points), y=points)) +
  geom_bar(stat="identity", fill=colour_list)+
  geom_text(aes(label=points), vjust=1.6, color="grey", size=3.5)+
  theme_minimal()+labs(x = 'FP3', y = 'Points')
```

#QUALY Representation
```{r}
#2021
points = c(64, 205, 16, 0, 0, 0, 153, 176, 44, 182, 9, 41, 31, 398, 430, 0, 129, 21, 315, 8, 0)
point_frame = data.frame(points,driv_short,colour_list)

quali <-ggplot(data=point_frame, aes(x=reorder(driv_short,points), y=points)) +
  geom_bar(stat="identity", fill=colour_list)+
  geom_text(aes(label=points), vjust=1.6, color="grey", size=3.5)+
  theme_minimal()+labs(x = 'QUALY', y = 'Points')
```

```{r}
plot <- grid.arrange(fp1, fp2, fp3, quali, nrow = 2, ncol=2)
```

#Car diferences in Qualy
```{r}
qualy_data = read.csv('q_dist.csv',sep = ',', dec = '.')
qualy_data[,3] = qualy_data[,3]/1000

p<-ggplot(qualy_data, aes(x=GP_ID, y=Distance, colour=Team)) +
  geom_line()+
  geom_point()+
    scale_x_discrete(limits = c('Bahrein', 'Imola', 'Portugal', 'España', 'Monaco', 'Azerbaiyan','Francia', 'Austria1', 'Austria2',  'GB', 'Hungria', 'Belgica', 'Holanda', 'Italia', 'Rusia', 'Turquia','EEUU', 'Mexico', 'Brasil', 'Qatar', 'Saudi',  'AbuDhabi'))+xlab('Circuit')+ylab('Diference to pole in seconds')+scale_y_continuous(breaks=seq(0, 8, 0.5))



p+scale_color_manual(values=c(ar,at,alp,am,fer,'grey',mcl,mer,rb,wil))

```

#PCA and Clustering
```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
